One-shot learning: From domain knowledge to action models

### 2019-02-22 doc/ijcai-19-deductivelearning/deductivelearning.tex
Intro completed

The learning of action models in planning has been typically addressed with inductive learning data-intensive approaches. From the pioneer learning system \ARMS~\cite{yang2007learning} to other more recent ones \cite{MouraoZPS12,zhuo2013action,kuvcera2018louga}, all of them require of the order of thousands of actions (input plan observations) to obtain and validate an action model. Generally speaking, the rationale behind these approaches is to obtain the statistically significant model $M$ that best explains the plan observations by minimizing some error and redundancy metrics. Specifically, given a plan observation $o=\langle s_0,a_1,s_1,\ldots,s_{n-1},a_n,s_n\rangle$, being $s_i, i>0$, a possibly partially observable state; if a solution plan $\pi$ to the planning task $\tup{s_0,s_n}$ can be produced with $M$ such that $\pi$ contains the observed actions $\tup{a_1,\ldots,a_n}$ of $o$, and the trace resulting from the execution of $\pi$ comprises the (partially) observed states $\tup{s_1,\ldots,s_{n-1}}$ of $o$, then $M$ explains $o$. Validating the explainability of a model as an optimization task over a test set of observations neither guarantees completeness (the model may not explain all the observations) nor correctness (a state that results from the execution of $\pi$ may contain contradictory information even though the state comprises the observed state).

Differently, other approaches rely on a symbolic-via learning. The Simultaneous Learning and Filtering (\SLAF) approach~\cite{AmirC08} exploits logical inference and builds a complete explanation of observations through a CNF formula that represents the initial belief state, and an observation $o$ that contains partially observable states. The formula is updated with every action and state of $o$, thus representing all possible transition relations consistent with $o$. \SLAF extracts all satisfying models of the learned formula with a SAT solver although the algorithm cannot effectively learn the preconditions of actions. A more recent approach addresses the learning of action models from $o$ as a planning task that searches the space of the all possible action models~\cite{aineto2018learning}. A plan here is conceived as a series of steps that determine the preconditions and effects of the model actions plus other steps that validate the formed actions in $o$. The advantage of this approach is that only requires about a total of 50 actions in the set $o$.

After this brief analysis, we raise the following question: can an action model be learnt from \emph{no observations} but using domain-specific knowledge?. The question is motivated by (a) the assumption that obtaining enough training observations is often difficult and costly, if not impossible in some domains~\cite{Zhuo15}; (b) we may not know the physics of the real-world domain being modeled but we may know certain pieces of knowledge about the domain; and (c) obtaining correct action models that are usable beyond their applicability to a set of testing observations. To this end, we opted for checking our hypothesis in the framework proposed in ~\cite{aineto2018learning} since this planning-based satisfiability approach allows us to configure additional constraints in the compilation scheme, it is able to work under a minimal set of observations and uses an off-the-shelf planner\footnote{We thank authors for providing us with the source files of their learning system.}. Ultimately, we aim to compare the informational power of domain observations against the representational power of domain-specific knowledge.

The paper is organized as follows ...

>>>>

The learning of action models in planning has been typically addressed with inductive learning data-intensive approaches. From the pioneer learning system \ARMS~\cite{yang2007learning} to more recent ones \cite{MouraoZPS12,zhuo2013action,kuvcera2018louga}, all of them require thousands of plan observations or training samples, i.e., sequences of actions as evidence of the execution of an observed agent, to obtain and validate an action model. These approaches return the statistically significant model that best explains the plan observations by minimizing some error metric. A model explains an observation if a plan containing the observed actions can be generated with the model and the states induced by this plan also include the possibly partially observed states. The limitation of posing model validation as an optimization task over a testing set of observations is that it neither guarantees completeness (the model may not explain all the observations) nor correctness (the states induced by the execution of the plan generated with the model may contain contradictory information).

Differently, other approaches rely on symbolic-via learning. The Simultaneous Learning and Filtering (\SLAF) approach~\cite{AmirC08} exploits logical inference and builds a complete explanation through a CNF formula that represents the initial belief state, and a plan observation. The formula is updated with every action and state of the observation, thus representing all possible transition relations consistent with it. \SLAF extracts all satisfying models of the learned formula with a SAT solver although the algorithm cannot effectively learn the preconditions of actions. A more recent approach addresses the learning of action models from plan observations as a planning task which searches the space of all possible action models~\cite{aineto2018learning}. A plan here is conceived as a series of steps that determine the preconditions and effects of the action models plus other steps that validate the formed actions in the observations. The advantage of this approach is that it only requires input samples of about a total of 50 actions.

This paper studies the impact of using mixed input data, i.e, automatically-collected plan observations and human-encoded domain-specific knowledge, in the learning of action models. Particularly, we aim to stress the extreme case of having a single observation sample and answer the question to whether the lack of training samples can be overcome with the supply of domain knowledge. The question is motivated by (a) the assumption that obtaining enough training observations is often difficult and costly, if not impossible in some domains~\cite{Zhuo15}; (b) the fact that although the physics of the real-world domain being modeled are unknown, the user may know certain pieces of knowledge about the domain; and (c) the desire for correct action models that are usable beyond their applicability to a set of testing observations. To this end, we opted for checking our hypothesis in the framework proposed in~\cite{aineto2018learning} since this planning-based satisfiability approach allows us to configure additional constraints in the compilation scheme, it is able to work under a minimal set of observations and uses an off-the-shelf planner\footnote{We thank authors for providing us with the source files of their learning system.}. Ultimately, we aim to compare the informational power of domain observations (information quantity) with the representational power of domain-specific knowledge (information quality). Complementarily, we restrict our attention to solely observations over fluents as in many applications the actual actions of an agent may not be observable~\cite{SohrabiRU16}.

Next section summarizes basic planning concepts and outlines the baseline learning approach~\cite{aineto2018learning}. Then we formalize our one-shot learning task with domain knowledge and subsequently we explain the task-solving process. Section 5 presents the experimental evaluation and last section concludes.
